---
title: "BribeR Data Guide"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{BribeR Data Guide}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment  = "#>",
  fig.width = 7,
  fig.height = 4
)

```

## Overview

BribeR bundles seven datasets that together describe the Vladivideos
corpus from different angles. This article explains what each dataset
contains, which columns they share, and how to combine them for
analysis.

## The Seven Datasets

### `compiled_transcripts` — Full Corpus

The main dataset. Each row is one speech turn within a transcript.

```{r compiled, message=FALSE}
library(BribeR)
library(dplyr)

transcripts <- read_transcripts()
glimpse(transcripts)
```

| Column        | Type      | Description                          |
|---------------|-----------|--------------------------------------|
| `n`           | numeric   | Transcript identifier                |
| `row_id`      | numeric   | Row number within the transcript     |
| `date`        | character | Recording date                       |
| `speaker`     | character | Raw speaker label from the source    |
| `speech`      | character | Speech text (Spanish)                |
| `speaker_std` | character | Standardized speaker identifier      |
| `topic`       | character | Primary topic tag for the transcript |

The `speaker_std` column is the key linking column across all datasets.
It uses uppercase surnames and is consistent with the `actors`,
`speakers_per_transcript`, and `transcript_index` datasets.

### `descriptions` — Transcript Metadata

One row per transcript. Contains dates, topic flags, availability
information, and plain-language English summaries.

```{r descriptions}
head(descriptions[, c("n", "date", "type", "summary")])
```

The `topic_*` columns use `"x"` to indicate a topic is present:

```{r topic-flags}
names(descriptions)[grepl("^topic_", names(descriptions))]
```

### `transcript_index` — Fast Lookup Table

A wide-format boolean table with one row per transcript and one column
per speaker and per topic. Designed for fast filtering without loading
the full corpus.

```{r index}
dim(transcript_index)
```

The `speaker_*` and `topic_*` columns contain `1` (present) or `0`
(absent). This is the dataset used internally by `get_transcript_id()`.

### `speakers_per_transcript` — Speaker Roster

One row per transcript, with up to 19 speaker slots as separate columns.

```{r speakers-wide}
speakers_per_transcript[1, ]
```

Note: the column names contain a known typo (`speakrer` instead of
`speaker`) preserved from the source data. The package handles this
automatically in all functions.

### `actors` — Actor Roster

Biographical and institutional metadata for 125 individuals named in the
transcripts.

```{r actors}
head(actors[, c("speaker", "Position", "Type", "speaker_std")])
```

The `Type` column classifies actors into institutional categories:
`Security`, `Congress`, `Judiciary`, `Media`, `Businessperson`,
`Elected Official`, `Bureaucrat`, `Foreign`, and `Illicit`.

The `speaker_std` column links to `compiled_transcripts`,
`transcript_index`, and `speakers_per_transcript`.

### `actors_description` — Actor Descriptions

Short biographical descriptions for 79 actors, keyed by `speaker_std`.

```{r actor-desc}
head(actors_description)
```

### `topic_descriptions` — Topic Reference

Human-readable labels and descriptions for each of the 15 topic tags.

```{r topic-desc}
topic_descriptions
```

## Linking Datasets

The table below shows which columns connect the datasets:

| From                   | To                        | Key column         |
|------------------------|---------------------------|--------------------|
| `compiled_transcripts` | `actors`                  | `speaker_std`      |
| `compiled_transcripts` | `transcript_index`        | `n`                |
| `compiled_transcripts` | `speakers_per_transcript` | `n`                |
| `compiled_transcripts` | `descriptions`            | `n`                |
| `actors`               | `actors_description`      | `speaker_std`      |
| `descriptions`         | `topic_descriptions`      | topic column names |

## Using `speaker_std` as a Linking Key

The `speaker_std` column is the consistent identifier linking actors
across all datasets. It uses uppercase surnames and resolves naming
variation in the original source material. When writing analysis code,
always use `speaker_std` rather than the raw `speaker` column for joins
and filters.

```{r speaker-std-example}
transcripts |>
  filter(speaker_std == "MONTESINOS") |>
  summarise(n_turns = n(), n_words = sum(nchar(speech)))
```

`get_transcript_id()` expects lowercase values for the `speaker`
argument, which it matches against lowercase `speaker_std` values
internally:

```{r get-id-lowercase}
get_transcript_id(speaker = "montesinos")
```

## Accessing Data Directly

All datasets are lazily loaded when the package is attached, so you can
reference them by name after `library(BribeR)`:

```{r direct-access}
nrow(compiled_transcripts)
nrow(actors)
nrow(topic_descriptions)
```
