---
title: "Corpus Overview"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Corpus Overview}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment  = "#>",
  eval     = FALSE
)
```

## The Vladivideos

Between 1990 and 2000, Vladimiro Montesinos Torres — Peru's de facto
intelligence chief under President Alberto Fujimori — secretly recorded
thousands of meetings in which he bribed politicians, judges, military
officers, media executives, and businesspeople. When the recordings were
leaked and broadcast publicly in September 2000, they triggered the collapse
of the Fujimori government and became one of the most extensively documented
cases of systemic corruption in Latin American history.

The videos capture corruption across every major institution of the Peruvian
state: legislators accepting cash to switch party allegiances, television
channel owners receiving monthly payments to control news coverage, military
generals coordinating electoral fraud, and judges confirming their
availability to rule in Montesinos's favor.

**{BribeR}** provides structured access to transcripts of 101 of these
recordings, along with metadata on the speakers involved and the topics
discussed.

## Transcripts

```{r meta-load}
library(BribeR)
library(dplyr)

meta <- read_transcript_meta_data()
nrow(meta)
#> [1] 101
```

The corpus spans recordings made between 1996 and 2000, covering the period
from Fujimori's successful bid for a controversial third term through the
final months before the regime's collapse.

```{r word-counts}
# Distribution of transcript length
summary(meta$n_words)
#>    Min. 1st Qu.  Median    Mean 3rd Qu.    Max.
#>     102    1377    2502    3028    3956   18021
```

Transcripts range from brief exchanges of a few hundred words to lengthy
multi-hour meetings exceeding 18,000 words. The median transcript is
approximately 2,500 words.

## Actors

The corpus involves 125 named individuals. Each actor in the `actors` dataset
is classified by institutional type:

```{r actors-table}
actors |>
  count(Type, sort = TRUE)
#> # A tibble: 9 × 2
#>   Type                  n
#>   <chr>             <int>
#> 1 Security             31
#> 2 Congress             25
#> 3 Judiciary            18
#> 4 Media                14
#> 5 Businessperson       12
#> 6 Elected Official      9
#> 7 Bureaucrat            8
#> 8 Illicit               5
#> 9 Foreign               3
```

Security forces (military and police) and Congress form the two largest
groups, reflecting Montesinos's priorities: controlling the armed forces and
ensuring a legislative majority for Fujimori's agenda.

The most frequently appearing speaker across transcripts is Montesinos himself:

```{r speaker-freq}
speakers <- get_transcript_speakers()

speakers |>
  mutate(n_transcripts = lengths(transcripts)) |>
  arrange(desc(n_transcripts)) |>
  head(10)
#> # A tibble: 10 × 3
#>    speaker_std        transcripts n_transcripts
#>    <chr>              <list>              <int>
#>  1 MONTESINOS         <dbl [98]>             98
#>  2 CROUSILLAT         <dbl [12]>             12
#>  3 KOURI              <dbl [10]>             10
#>  4 DELGADO PARKER     <dbl [9]>               9
#>  5 FUJIMORI           <dbl [8]>               8
#>  ...
```

## Topics

Each transcript is tagged with one or more thematic topics. The 15 topics
in the corpus reflect the main areas of Montesinos's corrupt activity:

| Topic | Description |
|---|---|
| `state_capture` | Systematic co-optation of state institutions |
| `media` | Payments to television channels and newspapers |
| `reelection` | Planning and executing Fujimori's 2000 reelection campaign |
| `congress` | Bribing legislators to switch parties or support legislation |
| `judiciary` | Influencing judges and prosecutors |
| `promotions` | Military and police promotions in exchange for loyalty |
| `foreign` | Foreign policy and international relations |
| `ivcher` | The case of Baruch Ivcher, a media owner stripped of citizenship |
| `canal4` | Dealings with Canal 4 (RBC Televisión) |
| `safety` | Internal security and anti-opposition operations |
| `state_capture` | Cross-institutional coordination of authoritarian control |
| `wiese` | The Wiese banking group |
| `lucchetti_factory` | The Lucchetti factory zoning controversy |
| `ecuador` | The 1995 border conflict with Ecuador |
| `referendum` | The 1993 constitutional referendum |
| `municipal98` | Lima's 1998 municipal elections |
| `miraflores` | The Miraflores district municipality |
| `public_officials` | Interactions with civil servants and bureaucrats |

```{r topic-counts}
# Transcripts per topic
get_transcript_id(topic = "media")      |> length()
#> [1] 14
get_transcript_id(topic = "reelection") |> length()
#> [1] 18
get_transcript_id(topic = "state_capture") |> length()
#> [1] 22
```

State capture — the coordinated takeover of multiple institutions — is the
most common topic tag, appearing in 22 transcripts. Many transcripts are
tagged with multiple topics: a single meeting might involve media payments,
legislative maneuvering, and electoral fraud simultaneously.

## Data Structure

The full compiled corpus contains 47,375 speech turns across the 101
transcripts:

```{r corpus-size}
transcripts <- read_transcripts()
nrow(transcripts)
#> [1] 47375

# Columns available
names(transcripts)
#> [1] "n"           "row_id"      "date"        "speaker"
#> [5] "speech"      "speaker_std" "topic"
```

Each row represents one turn in the conversation. The `speech` column
contains the original Spanish text; `speaker_std` is a standardized
uppercase surname identifier that matches across the `actors`,
`transcript_index`, and `speakers_per_transcript` datasets.

## Further Reading

- Conaghan, C. M. (2005). *Fujimori's Peru: Deception in the Public Sphere*.
  University of Pittsburgh Press.
- McMillan, J., & Zoido, P. (2004). How to subvert democracy: Montesinos in
  Peru. *Journal of Economic Perspectives*, 18(4), 69–92.