---
title: "Data Structure"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Data Structure}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment  = "#>",
  eval     = FALSE
)
```

## Overview

BribeR bundles seven datasets that together describe the Vladivideos corpus
from different angles. This article explains what each dataset contains,
which columns they share, and how to combine them for analysis.

## The Seven Datasets

### `compiled_transcripts` — Full Corpus

The main dataset. Each row is one speech turn within a transcript.

```{r compiled}
library(BribeR)

transcripts <- read_transcripts()
glimpse(transcripts)
#> Rows: 47,375
#> Columns: 7
#> $ n           <dbl>  1, 1, 1, 1, 1, ...
#> $ row_id      <dbl>  1, 2, 3, 4, 5, ...
#> $ date        <chr>  "3/25/1997", "3/25/1997", ...
#> $ speaker     <chr>  "BACKGROUND", "ALVA", "LEWIS", ...
#> $ speech      <chr>  "Declaraciones de...", "Yo en realidad...", ...
#> $ speaker_std <chr>  "BACKGROUND", "ALVA", "LEWIS", ...
#> $ topic       <chr>  "topic_foreign", "topic_foreign", ...
```

| Column | Type | Description |
|---|---|---|
| `n` | numeric | Transcript identifier |
| `row_id` | numeric | Row number within the transcript |
| `date` | character | Recording date |
| `speaker` | character | Raw speaker label from the source |
| `speech` | character | Speech text (Spanish) |
| `speaker_std` | character | Standardized speaker identifier |
| `topic` | character | Primary topic tag for the transcript |

The `speaker_std` column is the key linking column across all datasets.
It uses uppercase surnames and is consistent with the `actors`,
`speakers_per_transcript`, and `transcript_index` datasets.

### `descriptions` — Transcript Metadata

One row per transcript. Contains dates, topic flags, availability
information, and plain-language English summaries.

```{r descriptions}
head(descriptions[, c("n", "date", "type", "summary")])
#> # A tibble: 6 × 4
#>       n date      type  summary
#>   <dbl> <chr>     <chr> <chr>
#> 1   104 7/1/2000  audio "Montesinos convenes a meeting..."
#> 2   103 6/27/2000 audio "..."
#> ...
```

The `topic_*` columns use `"x"` to indicate a topic is present:

```{r topic-flags}
names(descriptions)[grepl("^topic_", names(descriptions))]
#>  [1] "topic_referendum"       "topic_ecuador"
#>  [3] "topic_lucchetti_factory" "topic_municipal98"
#>  [5] "topic_reelection"       "topic_miraflores"
#>  [7] "topic_canal4"           "topic_media"
#>  [9] "topic_promotions"       "topic_ivcher"
#> [11] "topic_foreign"          "topic_wiese"
#> [13] "topic_public_officials" "topic_safety"
#> [15] "topic_state_capture"
```

### `transcript_index` — Fast Lookup Table

A wide-format boolean table with one row per transcript and one column per
speaker and per topic. Designed for fast filtering without loading the full
corpus.

```{r index}
dim(transcript_index)
#> [1] 101 134

# Columns: n, file, format, date, topic_*, speaker_*, topic_count, speaker_count
```

The `speaker_*` and `topic_*` columns contain `1` (present) or `0` (absent).
This is the dataset used internally by `get_transcript_id()`.

### `speakers_per_transcript` — Speaker Roster

One row per transcript, with up to 19 speaker slots as separate columns.

```{r speakers-wide}
speakers_per_transcript[1, ]
#> # A tibble: 1 × 20
#>       n speakrer_std_1 speakrer_std_2 speakrer_std_3 speakrer_std_4 ...
#>   <dbl> <chr>          <chr>          <chr>          <chr>
#> 1     1 ALVA           LEWIS          BURNET         GARCIA         NA
```

Note: the column names contain a known typo (`speakrer` instead of
`speaker`) preserved from the source data. The package handles this
automatically in all functions.

### `actors` — Actor Roster

Biographical and institutional metadata for 125 individuals named in
the transcripts.

```{r actors}
head(actors[, c("speaker", "Position", "Type", "speaker_std")])
#> # A tibble: 6 × 4
#>   speaker              Position                          Type       speaker_std
#>   <chr>                <chr>                             <chr>      <chr>
#> 1 Vladimir Montesinos  Alberto Fujimori's Chief of Staff Security   MONTESINOS
#> 2 ...
```

The `Type` column classifies actors into institutional categories:
`Security`, `Congress`, `Judiciary`, `Media`, `Businessperson`,
`Elected Official`, `Bureaucrat`, `Foreign`, and `Illicit`.

The `speaker_std` column links to `compiled_transcripts`, `transcript_index`,
and `speakers_per_transcript`.

### `actors_description` — Actor Descriptions

Short biographical descriptions for 79 actors, keyed by `speaker_std`.

```{r actor-desc}
head(actors_description)
#> # A tibble: 6 × 2
#>   speaker_std description
#>   <chr>       <chr>
#> 1 MONTESINOS  "Fujimori's Chief of Staff"
#> ...
```

### `topic_descriptions` — Topic Reference

Human-readable labels and descriptions for each of the 15 topic tags.

```{r topic-desc}
topic_descriptions
#> # A tibble: 15 × 2
#>    topics                  descriptions
#>    <chr>                   <chr>
#>  1 topic_referendum        "Referendum to Peruvian constitution..."
#>  2 topic_ecuador           "..."
#> ...
```

## Linking Datasets

The table below shows which columns connect the datasets:

| From | To | Key column |
|---|---|---|
| `compiled_transcripts` | `actors` | `speaker_std` |
| `compiled_transcripts` | `transcript_index` | `n` |
| `compiled_transcripts` | `speakers_per_transcript` | `n` |
| `compiled_transcripts` | `descriptions` | `n` |
| `actors` | `actors_description` | `speaker_std` |
| `descriptions` | `topic_descriptions` | topic column names |

## Example: Joining Transcripts with Actor Metadata

```{r join-example}
library(dplyr)

transcripts <- read_transcripts()

transcripts |>
  left_join(
    actors |> select(speaker_std, Type, Position),
    by = "speaker_std"
  ) |>
  filter(!is.na(Type)) |>
  count(Type, sort = TRUE)
#> # A tibble: 9 × 2
#>   Type                  n
#>   <chr>             <int>
#> 1 Security          15823
#> 2 Congress           8741
#> 3 Judiciary          4302
#> ...
```

## Accessing Data Directly

All datasets are lazily loaded when the package is attached, so you can
reference them by name after `library(BribeR)`:

```{r direct-access}
library(BribeR)

# Access datasets directly
nrow(compiled_transcripts)   # 47375
nrow(actors)                 # 125
nrow(topic_descriptions)     # 15
```