---
title: "Actors & Topics"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Actors & Topics}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment  = "#>",
  eval     = FALSE
)
```

## Overview

This article documents the actors and topics metadata bundled with BribeR.
Understanding who is in the corpus and what they were discussing is essential
for filtering, subsetting, and interpreting the transcripts.

## Actors

The `actors` dataset contains biographical and institutional metadata for
125 individuals named in the Vladivideos transcripts. Each person is
classified by their institutional role at the time of the recordings.

```{r actors-load}
library(BribeR)

head(actors[, c("speaker", "Position", "Type", "Party", "speaker_std")])
#> # A tibble: 6 × 5
#>   speaker              Position                          Type     Party speaker_std
#>   <chr>                <chr>                             <chr>    <chr> <chr>
#> 1 Vladimir Montesinos  Alberto Fujimori's Chief of Staff Security NA    MONTESINOS
#> 2 Alberto Fujimori     President of Peru                 Elected… C90   FUJIMORI
#> ...
```

### Institutional Types

Actors are grouped into nine institutional categories:

| Type | Count | Description |
|---|---|---|
| `Security` | 31 | Military officers and police commanders |
| `Congress` | 25 | Members of Congress, including opposition members bribed to switch allegiance |
| `Judiciary` | 18 | Judges, prosecutors, and members of the electoral tribunal |
| `Media` | 14 | Television channel owners and newspaper directors |
| `Businessperson` | 12 | Private sector executives and financiers |
| `Elected Official` | 9 | Mayors, regional presidents, and other elected officials |
| `Bureaucrat` | 8 | Senior civil servants and agency heads |
| `Illicit` | 5 | Individuals primarily associated with organized crime |
| `Foreign` | 3 | Foreign officials and diplomats |

### Finding Transcripts by Actor Type

You can combine the `actors` dataset with `get_transcript_speakers()` to
filter the corpus by institutional type:

```{r filter-by-type}
library(dplyr)

# Get all media-sector actors
media_actors <- actors |>
  filter(Type == "Media") |>
  pull(speaker_std) |>
  tolower()

# Find transcripts featuring any media actor
media_ids <- get_transcript_id(speaker = media_actors)
length(media_ids)
#> [1] 14
```

### Actor Descriptions

Short biographies are available for 79 actors in the `actors_description`
dataset, keyed by `speaker_std`:

```{r actor-desc}
actors_description |>
  filter(speaker_std == "MONTESINOS")
#> # A tibble: 1 × 2
#>   speaker_std description
#>   <chr>       <chr>
#> 1 MONTESINOS  "Fujimori's Chief of Staff"
```

To view a description for any speaker appearing in the transcripts:

```{r desc-join}
transcripts <- read_transcripts(transcripts = 1)

transcripts |>
  distinct(speaker_std) |>
  left_join(actors_description, by = "speaker_std")
```

## Topics

Each transcript is tagged with one or more of 15 thematic topics. Topics
reflect the main areas of Montesinos's corrupt activity captured in the
recordings.

```{r topics-list}
topic_descriptions
#> # A tibble: 15 × 2
#>    topics                   descriptions
#>    <chr>                    <chr>
#>  1 topic_referendum         "Referendum to Peruvian constitution..."
#>  2 topic_ecuador            "..."
#>  3 topic_lucchetti_factory  "..."
#>  ...
```

### Topic Reference

| Topic tag | Theme |
|---|---|
| `state_capture` | Systematic co-optation of state institutions across multiple sectors |
| `reelection` | Planning and executing Fujimori's contested 2000 reelection |
| `media` | Payments to television channels and newspapers for favorable coverage |
| `congress` | Bribing legislators to switch parties or support key legislation |
| `judiciary` | Influencing judges, prosecutors, and the electoral tribunal |
| `promotions` | Military and police promotions granted in exchange for political loyalty |
| `safety` | Internal security operations against the political opposition |
| `foreign` | Foreign policy, international relations, and the 1995 Ecuador border conflict |
| `ivcher` | The case of Baruch Ivcher, a media owner stripped of Peruvian citizenship |
| `canal4` | Dealings with Canal 4 (RBC Televisión) and its ownership |
| `wiese` | Interactions involving the Wiese banking group |
| `lucchetti_factory` | The Lucchetti pasta factory zoning controversy in Lima |
| `referendum` | The 1993 constitutional referendum supporting Fujimori's agenda |
| `municipal98` | Lima's 1998 municipal elections and political manipulation |
| `miraflores` | Dealings involving the Miraflores district municipality |
| `public_officials` | Interactions with civil servants and administrative officials |
| `ecuador` | Diplomatic and military dimensions of the Peru–Ecuador border conflict |

### Finding Transcripts by Topic

```{r find-by-topic}
# Single topic
state_capture_ids <- get_transcript_id(topic = "state_capture")
length(state_capture_ids)
#> [1] 22

# Multiple topics (OR logic)
media_reelection_ids <- get_transcript_id(topic = c("media", "reelection"))

# Combine speaker and topic filters
montesinos_media <- get_transcript_id(
  speaker = "montesinos",
  topic   = "media"
)
```

### Topic Co-occurrence

Many transcripts cover multiple topics simultaneously — a single meeting
might involve media payments, electoral fraud, and legislative manipulation
at once. You can explore this using `read_transcript_meta_data()`:

```{r co-occurrence}
library(dplyr)

meta <- read_transcript_meta_data()

# How many topics does each transcript cover?
meta |>
  mutate(n_topics = lengths(topics)) |>
  count(n_topics, sort = TRUE)
#> # A tibble: 5 × 2
#>   n_topics     n
#>      <int> <int>
#> 1        1    58
#> 2        2    29
#> 3        3    10
#> 4        0     3
#> 5        4     1

# Transcripts covering the most topics
meta |>
  mutate(n_topics = lengths(topics)) |>
  arrange(desc(n_topics)) |>
  select(n, date, n_topics, topics) |>
  head(5)
```

## Using `speaker_std` as a Linking Key

The `speaker_std` column is the consistent identifier linking actors across
all datasets. It uses uppercase surnames and resolves naming variation in
the original source material. When writing analysis code, always use
`speaker_std` rather than the raw `speaker` column for joins and filters.

```{r speaker-std-example}
# Correct: use speaker_std for filtering
transcripts |>
  filter(speaker_std == "MONTESINOS") |>
  summarise(n_turns = n(), n_words = sum(nchar(speech)))

# get_transcript_id() also expects lowercase speaker_std values
get_transcript_id(speaker = "montesinos")   # lowercase
```