---
title: "Getting Started with BribeR"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started with BribeR}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment  = "#>",
  fig.width = 7,
  fig.height = 4
)
```

## Overview

**{BribeR}** is an R package for accessing and analyzing the *Vladivideos* —
a collection of secret recordings documenting Vladimiro Montesinos, Peru's
intelligence chief under President Alberto Fujimori, bribing politicians,
judges, military officers, media owners, and businesspeople throughout the
1990s. The recordings were declassified and made public in 2000 and have
since become a landmark dataset for the study of corruption, state capture,
and authoritarian politics.

The package provides structured, machine-readable access to 101 transcripts
along with metadata on speakers, topics, and institutional roles. It includes
six main functions:

| Function | Description |
|---|---|
| `read_transcripts()` | Load the full transcript corpus, optionally filtered by ID |
| `get_transcript_id()` | Find transcript IDs by speaker or topic |
| `get_transcript_speakers()` | Get speakers associated with each transcript |
| `get_transcripts_raw()` | Load the raw source CSV files |
| `read_transcript_meta_data()` | Build a tidy metadata summary per transcript |
| `run_transcript_network_app()` | Launch an interactive network visualization app |

## Installation

```{r install, eval=FALSE}
# Install from CRAN
install.packages("BribeR")

# Or install the development version from GitHub
remotes::install_github("jessietrudeau/BribeR")
```

## Loading Transcripts

The main entry point is `read_transcripts()`, which loads the full corpus as
a tidy data frame. Each row is one speech turn, with columns for the
transcript ID (`n`), speaker, speech text, standardized speaker name, date,
and topic.

```{r read-all}
library(BribeR)

transcripts <- read_transcripts()
head(transcripts)
```

You can filter to one or more transcripts by passing their numeric IDs:

```{r read-subset}
t1     <- read_transcripts(transcripts = 1)
subset <- read_transcripts(transcripts = c(5, 12, 47))

nrow(t1)
nrow(subset)
```

## Finding Transcripts by Speaker or Topic

`get_transcript_id()` returns transcript IDs matching any combination of
speakers and topics. Filters use OR logic: a transcript is returned if
**any** of the specified speakers appear in it, or if **any** of the
specified topics are flagged.

```{r get-ids}
# All available transcript IDs
all_ids <- get_transcript_id()
length(all_ids)

# Transcripts featuring Montesinos
montesinos_ids <- get_transcript_id(speaker = "montesinos")
length(montesinos_ids)

# Transcripts about media manipulation
media_ids <- get_transcript_id(topic = "media")
length(media_ids)
```

Available topic names include: `referendum`, `ecuador`, `lucchetti_factory`,
`municipal98`, `reelection`, `miraflores`, `canal4`, `media`, `promotions`,
`ivcher`, `foreign`, `wiese`, `public_officials`, `safety`, and
`state_capture`.

## Getting Speakers Per Transcript

`get_transcript_speakers()` returns a tibble with one row per speaker and a
list-column of the transcript IDs they appear in.

```{r get-speakers}
speakers <- get_transcript_speakers()
head(speakers)
```

Each entry in the `transcripts` column is a numeric vector of IDs. You can
use `lengths()` to count appearances:

```{r speaker-counts}
library(dplyr)

speakers |>
  mutate(n_transcripts = lengths(transcripts)) |>
  arrange(desc(n_transcripts)) |>
  head(10)
```

## Transcript Metadata

`read_transcript_meta_data()` assembles a one-row-per-transcript summary
combining dates, speaker rosters, word counts, and topic flags.

```{r metadata}
meta <- read_transcript_meta_data()
head(meta)
```

```{r metadata-work}
# Five longest transcripts by word count
meta |>
  arrange(desc(n_words)) |>
  select(n, date, n_words) |>
  head(5)
```

## Loading Raw Transcripts

`get_transcripts_raw()` provides access to the original source CSV files,
useful when you need the unprocessed data before compilation.

```{r raw, eval=FALSE}
# Load transcript 3 as a data frame
t3 <- get_transcripts_raw(n = 3)

# Load multiple transcripts combined into a single tibble
combined <- get_transcripts_raw(n = c(3, 19, 47), combine = TRUE)
```

## Putting It Together

A common workflow is to identify transcripts of interest, load them, and
then analyze the text. The example below finds all transcripts about media
manipulation, loads them, and summarizes how much each speaker said.

```{r workflow}
library(ggplot2)

# Step 1: find relevant transcript IDs
media_ids <- get_transcript_id(topic = "media")

# Step 2: load just those transcripts
media_transcripts <- read_transcripts(transcripts = media_ids)

# Step 3: count words spoken per actor
media_transcripts |>
  mutate(n_words = lengths(strsplit(speech, "\\s+"))) |>
  group_by(speaker_std) |>
  summarise(total_words = sum(n_words, na.rm = TRUE), .groups = "drop") |>
  arrange(desc(total_words)) |>
  head(10) |>
  ggplot(aes(x = reorder(speaker_std, total_words), y = total_words)) +
  geom_col(fill = "#8B1A1A") +
  coord_flip() +
  labs(
    title = "Words spoken in media-related transcripts",
    x     = NULL,
    y     = "Total words"
  ) +
  theme_minimal(base_size = 13)
```

## Network Visualization App

`run_transcript_network_app()` launches an interactive Shiny application
with two network views:

- **Speaker–Topic Network**: connects speakers to the topics discussed in
  their transcripts. Node size reflects the number of transcripts a speaker
  appears in; color reflects institutional type.
- **Speaker Co-Appearance Network**: connects speakers who appear in the
  same transcript, with edge weight proportional to co-appearance frequency.

```{r app, eval=FALSE}
run_transcript_network_app()
```