---
title: "Getting Started with BribeR"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started with BribeR}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment  = "#>",
  eval     = FALSE
)
```

## Overview

**{BribeR}** is an R package for accessing and analyzing the *Vladivideos* —
a collection of secret recordings documenting Vladimiro Montesinos, Peru's
intelligence chief under President Alberto Fujimori, bribing politicians,
judges, military officers, media owners, and businesspeople throughout the
1990s. The recordings were declassified and made public in 2000, and have
since become a landmark dataset for the study of corruption, state capture,
and authoritarian politics.

The package provides structured, machine-readable access to 101 transcripts
along with metadata on speakers, topics, and institutional roles. It includes
six main functions:

| Function | Description |
|---|---|
| `read_transcripts()` | Load the full transcript corpus, optionally filtered by ID |
| `get_transcript_id()` | Find transcript IDs by speaker or topic |
| `get_transcript_speakers()` | Get speakers associated with each transcript |
| `get_transcripts_raw()` | Load the raw source CSV files |
| `read_transcript_meta_data()` | Build a tidy metadata summary per transcript |
| `run_transcript_network_app()` | Launch an interactive network visualization app |

## Installation

```{r install}
# Install from CRAN
install.packages("BribeR")

# Or install the development version from GitHub
# remotes::install_github("jessietrudeau/BribeR")
```

## Loading Transcripts

The main entry point is `read_transcripts()`, which loads the full corpus as
a tidy data frame. Each row is one speech turn, with columns for the
transcript ID (`n`), speaker, speech text, standardized speaker name, date,
and topic.

```{r read-all}
library(BribeR)

transcripts <- read_transcripts()
head(transcripts)
#> # A tibble: 6 × 7
#>       n row_id date      speaker    speech                  speaker_std topic
#>   <dbl>  <dbl> <chr>     <chr>      <chr>                   <chr>       <chr>
#> 1     1      1 3/25/1997 BACKGROUND Declaraciones de…       BACKGROUND  topic_foreign
#> 2     1      2 3/25/1997 ALVA       Yo en realidad…         ALVA        topic_foreign
#> 3     1      3 3/25/1997 LEWIS      Podría usted…           LEWIS       topic_foreign
#> ...
```

You can filter to one or more transcripts by passing their numeric IDs:

```{r read-subset}
# Load a single transcript
t1 <- read_transcripts(transcripts = 1)

# Load a selection of transcripts
subset <- read_transcripts(transcripts = c(5, 12, 47))
```

## Finding Transcripts by Speaker or Topic

`get_transcript_id()` returns transcript IDs matching any combination of
speakers and topics. Filters use OR logic: a transcript is returned if
**any** of the specified speakers appear in it, or if **any** of the
specified topics are flagged.

```{r get-ids}
# All available transcript IDs
all_ids <- get_transcript_id()
length(all_ids)
#> [1] 101

# Transcripts featuring Montesinos
montesinos_ids <- get_transcript_id(speaker = "montesinos")

# Transcripts about media manipulation
media_ids <- get_transcript_id(topic = "media")

# Transcripts featuring Kouri OR about reelection
combined_ids <- get_transcript_id(speaker = "kouri", topic = "reelection")
```

Available topic names include: `referendum`, `ecuador`, `lucchetti_factory`,
`municipal98`, `reelection`, `miraflores`, `canal4`, `media`, `promotions`,
`ivcher`, `foreign`, `wiese`, `public_officials`, `safety`, and
`state_capture`.

## Getting Speakers Per Transcript

`get_transcript_speakers()` returns a tibble with one row per speaker and a
list-column of the transcript IDs they appear in. You can filter by
transcript ID, topic, or both.

```{r get-speakers}
# All speakers across all transcripts
speakers <- get_transcript_speakers()
head(speakers)
#> # A tibble: 6 × 2
#>   speaker_std transcripts
#>   <chr>       <list>
#> 1 ALVA        <dbl [1]>
#> 2 AMERICANO   <dbl [2]>
#> 3 AMOIN       <dbl [1]>
#> 4 ARANCIBIA   <dbl [3]>
#> 5 ARCE        <dbl [1]>
#> 6 AYBAR       <dbl [1]>

# Speakers who appear in transcripts about media
media_speakers <- get_transcript_speakers(topic = "media")
```

## Transcript Metadata

`read_transcript_meta_data()` assembles a one-row-per-transcript summary
combining dates, speaker rosters, word counts, and topic flags from the
bundled datasets.

```{r metadata}
meta <- read_transcript_meta_data()
head(meta)
#> # A tibble: 6 × 5
#>       n date      speakers  n_words topics
#>   <dbl> <chr>     <list>      <int> <list>
#> 1     1 3/25/1997 <chr [4]>    2819 <chr [1]>
#> 2     2 3/19/1997 <chr [2]>    1020 <chr [1]>
#> 3     3 6/4/1998  <chr [3]>    3277 <chr [1]>
#> ...
```

The `speakers` and `topics` columns are list-columns. You can unnest them or
work with them directly:

```{r metadata-work}
library(dplyr)

# Five longest transcripts by word count
meta |>
  arrange(desc(n_words)) |>
  select(n, date, n_words) |>
  head(5)
```

## Loading Raw Transcripts

`get_transcripts_raw()` provides access to the original source CSV files.
This is useful when you need the unprocessed data or want to work with
specific transcripts without loading the full compiled corpus.

```{r raw}
# Load transcript 3 as a data frame
t3 <- get_transcripts_raw(n = 3)

# Load multiple transcripts combined into a single tibble
combined <- get_transcripts_raw(n = c(3, 19, 47), combine = TRUE)
```

## Network Visualization App

`run_transcript_network_app()` launches an interactive Shiny application
with two network views:

- **Speaker–Topic Network**: connects speakers to the topics discussed in
  their transcripts. Node size reflects the number of transcripts a speaker
  appears in; color reflects institutional type (Congress, Judiciary, Media,
  Security, etc.).
- **Speaker Co-Appearance Network**: connects speakers who appear in the same
  transcript, with edge weight proportional to co-appearance frequency.

```{r app}
run_transcript_network_app()
```

Both networks support hover tooltips with speaker biographies and topic
descriptions, and a dropdown to highlight any individual speaker's
connections across the network.