---
title: "Raw Data Guide"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Raw Data Guide}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment  = "#>",
  fig.width = 7,
  fig.height = 4
)

```

# About the Vladivideos

(PICTURE)

Between 1990 and 2000, Vladimiro Montesinos Torres — Peru's de facto
intelligence chief under President Alberto Fujimori — secretly recorded
thousands of meetings in which he bribed politicians, judges, military
officers, media executives, and businesspeople. When the recordings were
leaked and broadcast publicly in September 2000, they triggered the
collapse of the Fujimori government and became one of the most
extensively documented cases of systemic corruption in Latin American
history.

The videos capture corruption across every major institution of the
Peruvian state: legislators accepting cash to switch party allegiances,
television channel owners receiving monthly payments to control news
coverage, military generals coordinating electoral fraud, and judges
confirming their availability to rule in Montesinos's favor.

**{BribeR}** provides structured access to transcripts of 101 of these
recordings, along with metadata on the speakers involved and the topics
discussed.


Understanding who is in the corpus and what they were discussing is
essential for filtering, subsetting, and interpreting the transcripts.
... This page shows the raw data... I demonstrate how to use the
functions later on (LINK).


(EXAMPLE HERE)


## Transcripts

```{r meta-load, message=FALSE}
library(BribeR)
library(dplyr)
library(ggplot2)

meta <- read_transcript_meta_data()
nrow(meta)
```

The corpus spans recordings made between 1996 and 2000, covering the period
from Fujimori's successful bid for a controversial third term through the
final months before the regime's collapse.

```{r word-counts}
summary(meta$n_words)
```

Transcripts range from brief exchanges of a few hundred words to lengthy
multi-hour meetings exceeding 18,000 words. The median transcript is
approximately 2,500 words.

```{r length-hist, fig.alt="Histogram of transcript word counts showing a right-skewed distribution centered around 2500 words"}
ggplot(meta, aes(x = n_words)) +
  geom_histogram(bins = 25, fill = "#8B1A1A", color = "white") +
  labs(
    title = "Distribution of transcript length",
    x     = "Words per transcript",
    y     = "Count"
  ) +
  theme_minimal(base_size = 13)
```

## Actors

The `actors` dataset contains biographical and institutional metadata
for 125 individuals named in the Vladivideos transcripts. Each person is
classified by their institutional role at the time of the recordings.

```{r actors-load, message=FALSE}
head(actors[, c("speaker", "Position", "Type", "Party", "speaker_std")])
```

### Actor Types

Actors are grouped into nine categories:

| Type | Count | Description |
|------------------------|------------------------|------------------------|
| `Security` | 31 | Military officers and police commanders |
| `Congress` | 25 | Members of Congress, including opposition members bribed to switch allegiance |
| `Judiciary` | 18 | Judges, prosecutors, and members of the electoral tribunal |
| `Media` | 14 | Television channel owners and newspaper directors |
| `Businessperson` | 12 | Private sector executives and financiers |
| `Elected Official` | 9 | Mayors, regional presidents, and other elected officials |
| `Bureaucrat` | 8 | Senior civil servants and agency heads |
| `Illicit` | 5 | Individuals primarily associated with organized crime |
| `Foreign` | 3 | Foreign officials and diplomats |

```{r type-bar, fig.alt="Bar chart of actor counts by institutional type"}
actors |>
  count(Type, sort = TRUE) |>
  ggplot(aes(x = reorder(Type, n), y = n)) +
  geom_col(fill = "#8B1A1A") +
  coord_flip() +
  labs(
    title = "Actors by institutional type",
    x     = NULL,
    y     = "Number of actors"
  ) +
  theme_minimal(base_size = 13)
```

- Very few are desconocido
- Party is able to be linked to VParty ...

## Topics

Each transcript is hand-coded[^1] and classified by topic(s). The 15
topics include:

[^1]: Each transcript was read 2-3 times by native Spanish-language
    speakers and classified as pertaining to one or more relevant
    topics.

### Topic Types

| Topic | Description |
|------------------------------------|------------------------------------|
| `state_capture` | Co-optation of state institutions |
| `reelection` | Fujimori's 2000 reelection campaign |
| `media` | Bribery of television channels and newspapers |
| `promotions` | Military and police promotions in exchange for loyalty |
| `foreign` | Foreign policy and international relations |
| `ivcher` | The case of Baruch Ivcher, a media owner stripped of citizenship |
| `canal4` | Dealings with Canal 4 (RBC Televisión) |
| `safety` | Domestic public security and anti-opposition suppression |
| `wiese` | Wiese banking group |
| `lucchetti_factory` | Lucchetti factory zoning controversy |
| `ecuador` | 1995 border conflict with Ecuador |
| `referendum` | Presidental term limits referendum |
| `municipal98` | 1998 municipal elections |
| `miraflores` | 1998 Miraflores district elections |
| `public_officials` | Interactions with civil servants and bureaucrats |

```{r topic-counts, fig.alt="Bar chart showing number of transcripts tagged with each topic, with state_capture and reelection as the most common"}
topic_names <- c(
  "state_capture", "reelection", "media", "promotions", "foreign",
  "ivcher", "canal4", "safety", "wiese", "lucchetti_factory",
  "ecuador", "referendum", "municipal98", "miraflores", "public_officials"
)

topic_counts <- sapply(topic_names, function(t) {
  length(get_transcript_id(topic = t))
})

data.frame(topic = topic_names, n = topic_counts) |>
  ggplot(aes(x = reorder(topic, n), y = n)) +
  geom_col(fill = "#8B1A1A") +
  coord_flip() +
  labs(
    title = "Transcripts per topic",
    x     = NULL,
    y     = "Number of transcripts"
  ) +
  theme_minimal(base_size = 12)
```

As demonstrated in M and Z - media is the most common topic tag. Many
transcripts are tagged with two or more topics.


```{r co-occurrence-bar, fig.alt="Bar chart showing how many transcripts cover 1, 2, 3, or 4 topics simultaneously"}
meta |>
  mutate(n_topics = lengths(topics)) |>
  count(n_topics) |>
  ggplot(aes(x = factor(n_topics), y = n)) +
  geom_col(fill = "#8B1A1A") +
  labs(
    title = "Topics per transcript",
    x     = "Number of topics",
    y     = "Number of transcripts"
  ) +
  theme_minimal(base_size = 13)
```

# Further Reading

- Conaghan, C. M. (2005). *Fujimori's Peru: Deception in the Public
  Sphere*. University of Pittsburgh Press.
- McMillan, J., & Zoido, P. (2004). How to subvert democracy: Montesinos
  in Peru. *Journal of Economic Perspectives*, 18(4), 69–92.
