---
title: "Transcript Network App"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Transcript Network App}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment  = "#>"
)
```

## Overview

`run_transcript_network_app()` launches an interactive Shiny application
visualizing the relationships between speakers and topics across the
Vladivideos transcript corpus. It provides two network views:

- **Speaker-Topic Network** -- connects speakers to the topics discussed in
  their transcripts.
- **Speaker Co-Appearance Network** -- connects speakers who appear in the
  same transcript.

On the package website, the app below runs live in your browser via
[shinylive](https://posit-dev.github.io/r-shinylive/) -- a WebAssembly build
of R that needs no Shiny server. It behaves the same as calling
`run_transcript_network_app()` in RStudio or VS Code: select a speaker from
the dropdown, drag nodes, zoom, and hover for details.

```{=html}
<iframe
  src="network-app/index.html"
  width="100%"
  height="950"
  style="border: 1px solid #d5d4c7; border-radius: 4px;"
  allowfullscreen>
</iframe>
```

First load takes a few seconds while the in-browser R runtime and packages
initialize.

```{r, eval=FALSE}
# To run the same app locally instead:
library(BribeR)
run_transcript_network_app()
```

If you're reading this vignette outside the package website (e.g. via
`browseVignettes()`), the embedded app above won't load -- run
`run_transcript_network_app()` directly instead.
