---
title: "Using BribeR"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Using BribeR}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment  = "#>",
  fig.width = 7,
  fig.height = 4
)

```


**BribeR** provides structured, full-text access to 99 *Vladivideo* transcripts, along with metadata about the conversations, speakers, and topics. This vignette introduces the three main families of functions (below) and provides user-friendly examples for how to use **BribeR.**

**1. Read transcripts**

**2. Find transcripts**

**3. Integrate with metadata** 

To begin, install and load the package: 

```{r install, eval=FALSE}
# Install from CRAN
install.packages("BribeR")

# Or install the development version from GitHub
remotes::install_github("jessietrudeau/BribeR")
```

```{r load}
# Load packages
library(BribeR)
library(dplyr)
```

# Read transcripts

Use `read_transcripts()` to access the corpus of full-text transcripts as a tidy data frame. Each row corresponds to one speech turn, indexed by transcript ID (`id`), date, standardized speaker name, and speaker. The column `speech` includes the full-text transcription of one turn. 

```{r read-all}
# Get full-text transcripts 
transcripts <- read_transcripts()

# Additional variables
colnames(transcripts)

head(transcripts)
```

You can filter to one or more transcripts using their numeric IDs:

```{r read-subset}
# Select first transcript
t1 <- read_transcripts(transcripts = 1)
nrow(t1)

# Select multiple transcripts 
t_sub <- read_transcripts(transcripts = c(5, 12, 47))
nrow(t_sub)
```

## Read raw transcript data

`get_transcripts_raw()` provides access to the original source CSV files for users who wish to access the data before compilation.

```{r raw, eval=F}
# Load transcript 3 as a data frame
t3 <- get_transcripts_raw(n = 3)

# Load multiple transcripts combined into a single tibble
combined <- get_transcripts_raw(n = c(3, 19, 47), combine = TRUE)
```


# Find transcripts

If users know the actor(s) or topic(s) they wish to focus on, but not the numeric `id`, use `get_transcript_id()` to filter transcripts by speaker or topic characteristics.

## By actor 

The `get_transcript_id()` function accepts lowercase string values for the standardized speaker's name (the `speaker_std` variable), and returns transcript IDs where these speakers are present. 

```{r get-ids}
# Transcripts featuring Montesinos
montesinos_ids <- get_transcript_id(speaker = "montesinos")
length(montesinos_ids)
montesinos_ids

# Transcripts featuring Alex Kouri
kouri_ids <- get_transcript_id(speaker = "alex kouri")
length(kouri_ids)
kouri_ids
```

There are 125 valid speaker IDs that the `speaker_std` variable can take on, listed in the `actors` dataset.


## By topic 

Similarly, `get_transcript_id()` function accepts lowercase string values for topic names, and returns transcript IDs where these topics are discussed. 

```{r}
# Transcripts about media manipulation
media_ids <- get_transcript_id(topic = "media")
length(media_ids)
media_ids

# Transcripts about BOTH media and reelection 
media_reelection_ids <- get_transcript_id(topic = c("media", "reelection"))
length(media_reelection_ids)

```

There are 15 valid topics, listed in the [Raw Data Guide](https://jessietrudeau.com/BribeR/articles/raw_data_guide.html#topics). 

## By both

Finally, users can filter transcripts by actor and topic. This function uses AND logic and returns transcript IDs where both the selected speaker(s) are present and topic is mentioned.

```{r}
# Find transcripts about ecuador  
ecuador_ids <- get_transcript_id(topic = "ecuador")
length(ecuador_ids)

# Find transcripts about ecuador where a Sendero Luminoso (armed group) leader is present 
ecuador_morote_ids <- get_transcript_id(topic = "ecuador", speaker = "morote")
length(ecuador_morote_ids)

```

# Integrate with metadata

Rich transcript-level and actor-level metadata is available in **BribeR**. 

The `read_transcript_meta_data()` function presents transcript-level data containing dates, summaries,[^1] speakers present, topics mentioned, and word counts. 

When left blank, it returns the metadata for every transcript in the corpus, but can also be filtered using the transcript `id` variable.

```{r metadata}
meta <- read_transcript_meta_data()
head(meta)

# Get metadata for transcript 5 - THIS DOESN'T WORK
read_transcript_meta_data(5)
```


# Examples 

The two below examples demonstrate how the **BribeR** functions and data can be used together. 


## Example 1: Who speaks about media manipulation? For how long? 

Say that we are interested in how Montesinos and his counterparts talk about a commonly discussed topic, media manipulation. First, we start by selecting the conversations about media manipulation and counting which actors speak the most during these conversations: 

```{r}
library(ggplot2)

# Step 1: find transcript IDs about media 
media_ids <- get_transcript_id(topic = "media")

# Step 2: load just those transcripts about the media 
media_transcripts <- read_transcripts(transcripts = media_ids)

# Step 3: count words spoken per actor
media_transcripts |>
  mutate(n_words = lengths(strsplit(speech, "\\s+"))) |>
  group_by(speaker_std) |>
  summarise(total_words = sum(n_words, na.rm = TRUE), .groups = "drop") |>
  arrange(desc(total_words)) |>
  ## drop the 'background' and 'desconocido' to focus on identifiable actors only
  filter(speaker_std != "desconocido" & speaker_std != "background") |>
  head(10) |>
  ggplot(aes(x = reorder(speaker_std, total_words), y = total_words, fill = speaker_std)) +
  geom_col() + coord_flip() + theme_minimal(base_size = 13) +
  theme(legend.position = "none") +
  labs(
    title = "Words spoken in media-related transcripts, \n top-10 most likely speakers",
    x     = NULL,
    y     = "Total words"
  )
```

Second, let's say we're interested in making some aggregate conclusions about the type of actor that's speaking during such conversations. This requires crossing the corpus of media transcripts with the actor-level metadata. 

```{r}
# Step 1: Left join media_transcripts with actor-level metadata from `actors`
media_transcripts <- media_transcripts |>
  left_join(actors |> select(speaker_std, type, position),
    by = "speaker_std"
  )

## Step 2: Count words spoken in media_transcripts and group by type, summarize
media_transcripts |>
  mutate(n_words = lengths(strsplit(speech, "\\s+"))) |>
  group_by(type) |>
  summarise(total_words = sum(n_words, na.rm = TRUE), .groups = "drop") |>
  arrange(desc(total_words)) |> 
    filter(is.na(type) == F) |> #Filter only to identified actors
  ggplot(aes(x = reorder(type, total_words), y = total_words, fill = type)) +
  geom_col() + coord_flip() + theme_minimal(base_size = 13) +
  theme(legend.position = "none") +
  labs(
    title = "Words spoken in media-related transcripts, by actor type",
    x     = NULL,
    y     = "Total words"
  ) 

```


## Example 2: Finding Transcripts by Speaker Type

You can combine `actors` and `get_transcript_id()` to filter the corpus by
type of actor rather than by individual name. The example below
finds all transcripts featuring any media-sector actor who appears in the
transcript index:

```{r type-filter}
# Step 1: find all media actors 
media_actors <- actors |>
  filter(type == "media") |>
  pull(speaker_std) 
media_actors


# Step 2: Filter for transcripts where media actors are present
## ERROR: e.g. VALENZUELA
#media_ids <- get_transcript_id(speaker = media_actors)
#length(media_ids)
```


[^1]: These summaries were written by undergraduate native-Spanish speakers. 

